#Set the default of echo to be true throughout the document:
```{r}
library(knitr)
opts_chunk$set(echo = TRUE)
```
#Load and process data
```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}

activity<-read.csv("activity.csv")
head(activity)
```
#What is the mean number of steps taken per day
Ignore the missing values in the dataset. Calculate the total number of setps taken per day.
```{r}
totalStepsByDay<-aggregate(steps~date, activity, sum)
```
Make a histogram of the total number of steps taken each day
```{r chunk1, echo=FALSE}
hist(totalStepsByDay$steps, xlab="Class of Total Number of Steps per day", ylab="Number of Days", main="Total Number of Steps taken each day")
```
```{r, echo=FALSE}
plot(hist(totalStepsByDay$steps, xlab="Class of Total Number of Steps per day", ylab="Number of Days", main="Total Number of Steps taken each day"))
```
  
Calculate the mean and median
```{r}
mean_raw<-mean(totalStepsByDay$steps)
mean_raw
```
```{r}
median_raw<-median(totalStepsByDay$steps)
median_raw
```
#What is the average daily activity pattern?
The average number of steps taken:
```{r}
averageStepsbyInterval<-aggregate(steps~interval, activity, mean)
```
Time series plot of the average number of steps taken:
```{r chunk2, echo=FALSE}
chunk2<-with(averageStepsbyInterval, plot(interval, steps, type = "l"))
```


Which 5-minute interval, on average, contains the maximum number of steps
```{r}
averageStepsbyInterval[which.max(averageStepsbyInterval[,2]),1]
```
#Imputing missing values
Calculate and report the total number of missing values
```{r}
missingIndex<-is.na(activity[,1])
```
Find the mean number of steps per interval
```{r}
m<-mean(averageStepsbyInterval$steps)
```
New dataset with missing data filled in
```{r}
activity1<-activity
activity1[missingIndex,1]<-m
head(activity1)
```
Histogram with missing data filled in
```{r}
totalStepsByDay1<-aggregate(steps~date, activity1, sum)
hist(totalStepsByDay1$steps, xlab="Class of Total Number of Steps per day", ylab="Number of Days", main="Number of Steps taken each day after missing values are imputed")
```
Calculate new mean and median
```{r}
totalStepsByDay1<-aggregate(steps~date, activity1, sum)
```
```{r}
mean_afterImput<-mean(totalStepsByDay1$steps)
mean_afterImput
```
```{r}
median_afterImput<-median(totalStepsByDay1$steps)
median_afterImput
```
#Are there differences in activity patterns between weekdays and weekends?
Create new factor variable with two levels - "weekday" and "weekend"
```{r}
activity1$date<-as.Date(activity1$date)
library(dplyr)
```
```{r}
activity2<-activity1%>%
        mutate(dayType= ifelse(weekdays(activity1$date)=="Saturday" | weekdays(activity1$date)=="Sunday", "Weekend", "Weekday"))
head(activity2)
```
Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
averageStepByDayTypeAndInterval<-activity2 %>%
  group_by(dayType, interval) %>%
  summarize(averageStepByDay=sum(steps))

head(averageStepByDayTypeAndInterval)
```
```{r}
library(lattice)
with(averageStepByDayTypeAndInterval, 
      xyplot(averageStepByDay ~ interval | dayType, 
      type = "l",      
      main = "Total Number of Steps within Intervals by dayType",
      xlab = "Daily Intervals",
      ylab = "Average Number of Steps"))
```

